Impact of grids and dynamical cores in CESM2.2 on the surface mass balance of the Greenland Ice Sheet

### 2022-04-05 arcticgrids.tex
bring big overleaf edits into github

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
\authors{Adam R. Herrington\affil{1}, Peter H. Lauritzen\affil{1}, William H. Lipscomb\affil{1}, Marcus Lofverstrom\affil{2} $\ $and Andrew Gettelman\affil{1}}

@@

\begin{keypoints}
\item enter point 1 here
\item enter point 2 here
\item enter point 3 here
\end{keypoints}

@@

\begin{abstract}
[ enter your Abstract here ]
\end{abstract}

@@

General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high-latitudes, which are among the most sensitive regions on Earth to global and environmental change. Despite their importance, the numerical treatment of polar regions in GCMs is handled in vastly-different ways due to the so-called \textit{pole-problem} \cite{W2007JMSJ}. The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a). Depending on the numerics, methods exist to suppress this instability, and latitude-longitude grids may be advantageous for polar processes as structures can be represented with more degrees of freedom than elsewhere in the computational domain. With the recent trend towards globally uniform unstructured grids, any potential benefits of latitude-longitude grids on polar regions may become a relic of the past. In this study a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), including brand new variable-resolution grids, are evaluated to understand their impacts on the simulated characteristics of the Arctic, with a special focus on the climate and surface mass balance of the Greenland Ice Sheet.

In the 1970's the pole problem was largely defeated through wide-spread adoption of efficient spectral transform methods in GCMs. These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g. horizontal derivatives) in the equation set can be solved for exactly. While spectral transform methods are still used in the 21st century, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems. The pole problem has thus re-emerged in contemporary climate models that use latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}. Polar filters are akin to a band-aid; they subdue the growth of unstable modes by applying additional damping to the solution over polar regions. This additional damping reduces the effective resolution in polar regions, and the resolved scales are approximately the same everywhere on the grid.

An alternative approach is to use unstructured grids, which allow for more flexible grid structures that permit quasi-uniform grid spacing globally and eliminates the pole-problem entirely (e.g., Figure~\ref{fig:uni-grids}c). This grid flexibility also permits variable-resolution or regional grid refinement (e.g., Figure~\ref{fig:vr-grids}). Grids can be developed with refinement over polar regions that could in principle make up for any loss in polar resolution in transitioning away from latitude-longitude grids (e.g., Figure~\ref{fig:vr-grids}), although this comes at the cost of a smaller CFL-limiting time-step in the refined region (the CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space). Unstructured grids also scale more efficiently on parallel systems than latitude-longitude grids, likely resulting in a greater prevalence of unstructured grids as computing power continued to increase over time.

The meteorology and climate of the Arctic is characterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}. For example, while synoptic scale storms are generally well represented at typical GCM resolutions of 1 to 2 degrees \cite{JW2006QJR,IPCCAR5}, mesoscale Polar Lows are not well resolved at these resolutions. These mesoscale systems are prevalent during the cold season and produce gale-force winds that can induce large heat and moisture fluxes through the underlying sea-ice/ocean interface. The Arctic also contains the Greenland Ice Sheet (hereafter denoted as \textit{GrIS}). While it blankets the largest island in the world (Greenland), many of the processes that control the GrIS annual surface mass balance (the integrated sum of precipitation and runoff) are only partially resolved at typical GCM resolutions. For example, GrIS precipitation is typically confined to the ice-sheet margins (predominately the southeastern margin) where orographic precipitation is generated by steep topographic slopes. GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically 10s to 100\,km wide and confined to low-level areas or regions with limited precipitation. GCMs struggle to resolve the magnitude and extent of these features \cite{VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite<e.g.,>{LETAL2020JAMES}.

The goal of this study is to characterize the representation of high-latitude regions using the spectral-element and finite-volume dycores in CESM2.2, as these models treat the high-latitudes, e.g., the pole-problem, in different ways. The manuscript is laid out as follows: Section~\ref{sec:methods} consists of documentation of the grids, dycores and physical parameterizations used in this study. The Arctic refined grids were developed by the authors, and this section serves as their official documentation in CESM2.2. Section~\ref{sec:methods} also contains a description of the experiments along with reanalysis datasets and post-processing software used for evaluating the model simulations. Section~\ref{sec:results} contains the results of the experiments, followed by Section~\ref{sec:conclusions} that provides a general discussion and conclusions.

@@

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=60mm]{figs/grid-ARCTIC.pdf}&
         \includegraphics[width=60mm]{figs/grid-ARCTICGRIS.pdf} \\
\end{tabular}
\end{center}
\caption{Spectral-element grid for the variable-resolution ARCTIC grid in this study. Note that this is not the computational grid; each element has $3\times3$ independent grid points.}
\label{fig:vr-grids}
\end{figure}

@@

The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 \cite<CAM;>{CAM63}, supports a number of different atmospheric dynamical cores. These include dycores using latitude-longitude grids, such as finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models. The EUL dycore is the oldest dycore in CAM, and the least supported of all the dycores. FV3 is the newest dycore in CAM, but it was not fully incorporated at the time this work commenced; both the EUL and FV3 dycores are omitted from this study. As such, the results presented in this study are comparing the performance of the SE and FV dycores.

@@

The FV dycore is a hydrostatic model that integrates the equations of motion using a finite-volume discretization on a spherical latitude-longitude grid \cite{LR1997QJR}. The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to Eulerian reference grid in the vertical \cite{L2004MWR}, using a hybrid-pressure vertical coordinate. Hyperviscous damping is applied to the divergent modes while Laplacian damping is applied to momentum in the top few layers, referred to as a \textit{sponge layer} \cite{L2011IJHPC}. A polar filter is used to avoid computational instability due to the convergence of the meridians, allowing for a more practical time-step. It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically in the poleward direction \cite{ST1995GEOS}.

@@

The SE dycore is a hydrostatic model that integrates the equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}. The computational domain is a cubed-sphere grid tiled with quadrilateral elements (e.g., Figure~\ref{fig:vr-grids}). Each element contains a fourth order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre (GLL) quadrature points. This results in 16 GLL nodal points within each element, with 12 of the points lying on the (shared) element boundary. Communication between elements happens via the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries that reconciles overlapping nodal values and produces a continuous global basis set.

As with the FV dycore, the dynamics evolve in floating Lagrangian layers that are subsequently mapped to an Eulerian reference grid. A dry mass vertical coordinate was more recently implemented for thermodynamic consistency with condensates \cite{LetAl2018JAMES}. The 2D dynamics have no implicit dissipation and so hyperviscosity operators are applied to all prognostic variables to remove spurious numerical errors \cite{DetAl2012IJHPCA}. Laplacian damping is applied in the sponge layer.

The SE dycore supports regional grid refinement via its variable-resolution configuration, requiring two enhancements over uniform resolution grids. (1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing. A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements. (2) The topography boundary conditions need to be smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{gmdd-8-4623-2015} has been modified to scale the smoothing radius by the local element size.

For spectral-element grids with quasi-uniform grid spacing, a variant in which tracer advection is computed using the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) is used instead \cite<>{LTOUNGK2017MWR}. CSLAM has improved tracer property preservation and accelerated multi-tracer transport. It uses a separate grid from the spectral-element dynamics, through dividing each element into $3\times3$ control volumes with quasi-equal area. The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the default SE dycore in which the physics are evaluated at the GLL nodal points \cite{HL2018MWR}.

@@

Six grid are evaluated in this study (Table~\ref{tbl:table1}). The FV dycore is run with $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $f09$ and $f19$, respectively (Figure~\ref{fig:uni-grids}a,b). The $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid is also run, referred to as $ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics. An additional $1^{\circ}$ CAM-SE-CSLAM grid is run, but with the physical paramerizations computed on a grid that contains $2\times2$ control volumes per element, $ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>{HETAL2019JAMES}.

Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic (Figure~\ref{fig:vr-grids}). The Arctic meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}). The $ARCTIC$ grid is a $1^{\circ}$ grid with $\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region. The $ARCTICGRIS$ grid is identical to the $ARCTIC$ grid, but contains an additional patch covering the big island of Greenland with $\frac{1}{8}^{\circ}$ resolution.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lccccc}
   \hline
   grid name & dycore & $\Delta x_{eq}$ (km) & $\Delta x_{refine}$ (km) & $\Delta t_{phys}$ (s) \\ 
   \hline
   $f19$ & FV & 278 & - &1800 \\
   $f09$ & FV & 139 & - &1800 \\
   $ne30pg2$ & SE-CSLAM & 167 & - & 1800 \\
   $ne30pg3$ & SE-CSLAM & 111 & - & 1800 \\
   $ARCTIC$ & SE & 111 & 28 & 450 \\
   $ARCTCIGRIS$ & SE & 111 & 14 & 225 \\
 \hline
 \end{tabular}
  \caption{Grids and dycores used in this study. $\Delta x_{eq}$ refers to average equatorial grid spacing, $\Delta x_{refine}$ refers to grid spacing in the refined region (if applicable) and $\Delta t_{phys}$ refers to the physics time-step. The dycore abbreviation FV refers to the finite-volume dycore, SE the spectral-element dycore and SE-CSLAM the spectral-element dycore w/ CSLAM tracer advection.}
 \label{tbl:table1}
 \end{table*}

@@

The CAM6 physical parameterization package (hereafter referred to as the \textit{physics}; \url{https://ncar.github.io/CAM/doc/build/html/index.html}) is used in all simulations in this study. CAM6 physics is most noteably different from it's predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a PBL, shallow convection and cloud macrophysics scheme. CLUBB is coupled with the MG2 microphysics scheme \cite{MG2}, with prognostic precipitation and classical nucleation theory in representing cloud ice for improved cloud-aerosol interactions. Deep convection is parameterized using a convective quasi-equilibrium mass flux scheme \cite{ZM1995AO,NRJ2008JC} inlcuding convective momentum transport \cite{RSG2010JAS}. PBL form drag is modeled after \cite{BBW2004QJRMS} and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale.

All grids and dycores in this study use 32 levels in the vertical, with a model top of about $1 \ hPa$ or about $40 \ km$. The physics time-step is dependent on grid resolution. Increases in horizontal resolution permit faster vertical velocities that reduce characteristic time-scales, and so the physics time-step is reduced to avoid large time truncation errors \cite{HR2018JAMES}. The $ARCTIC$ and $ARCTICGRIS$ grids are therefore run with a $4\times$ and $8\times$ reduction in physics time-step relative to the default 1800 s time-step used in coarser, uniform resolution grids (Table~\ref{tbl:table1}.

Initial simulations with the $ne30pg3$ spectral-element grid produced weaker shortwave cloud forcing relative to the tuned up finite-volume dycore. All runs with the spectral-element dycore have two CLUBB parameter changes in order to provide a more realistic cloud forcing and top-of-atmosphere radiation balance. These are CLUBB's $gamma$ parameter, reduced from 0.308 to 0.270, and c14, reduced from 2.2 to 1.6. Briefly, the $gamma$ parameter scales the width of the sub-grid distribution of vertical velocity, and c14 controls the strength of the damping term in the equation for the horizontal component of turbulent kinetic energy. For a thorough explanation of how CLUBB parameters impact the simulated climate, the reader is referred to \cite{GETAL2015JAMES}.

@@

All grids and dycores are run using an identical transient 1979-1998 AMIP-style configuration, with prescribed monthly SST/sea-ice after \cite{CESMSST}. This configuration refers to the $FHIST \ compset$ and runs out of the box in CESM2.2.

The surface mass balance (SMB) of the Greenland Ice Sheet (GrIS) is simulated in all grids and dycores in this study. The SMB is the sum of the mass source term, accumulation (i.e., precipitation), and the mass sink term, ablation. Ablation can be expressed as evaporation/sublimation plus total runoff, with runoff being a combination of liquid precipitation and snow and ice melt. Not all liquid precipitation becomes runs off the ice sheet; rain may penetrate pore spaces in the firn layer and freeze, forming ice lenses in the subsurface. These processes are represented by different components in CESM, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.

>>>>

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
\authors{Adam R. Herrington\affil{1}, Peter H. Lauritzen\affil{1}, William H. Lipscomb\affil{1}, Marcus Lofverstrom\affil{2} \ and Andrew Gettelman\affil{1} }

@@

\begin{keypoints}
\item The transition from latitude-longitude to unstructured grids in CESM2.2 adversely impacts the surface mass balance of the Greenland Ice Sheet.
\item Two Arctic refined mesh configurations have been developed and made available in CESM2.2.
\item The Arctic meshes substantially improves the surface mass balance over all $1^{\circ}$ grids, including the latitude-longitude grids.
\end{keypoints}

@@

\begin{abstract}

Six different configurations, a mixture of grids and atmospheric dynamical cores available in the Community Earth System Model, version 2.2 (CESM2.2), are evaluated for their skill in representing the climate of the Arctic and the surface mass balance of the Greenland Ice Sheet (GrIS). The more conventional $1^{\circ}-2^{\circ}$ uniform resolution grids systematically overestimate both accumulation and ablation over the GrIS. Of these conventional grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids owing to their greater degrees of freedom in representing the GrIS. Two Arctic-refined meshes, with $\frac{1}{4}^{\circ}$ and $\frac{1}{8}^{\circ}$ refinement over Greenland, are documented as newly supported configurations in CESM2.2. The Arctic meshes substantially improve the simulated clouds and precipitation rates in the Arctic. Over Greenland, these meshes skillfully represent accumulation and ablation processes, leading to a more realistic GrIS surface mass balance. As CESM is in the process of transitioning away from latitude-longitude grids, these new Arctic refined meshes improve the representation of polar processes in CESM by recovering resolution lost in the transition to quasi-uniform grids.

\end{abstract}

@@

General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high latitudes, which are among the most sensitive regions on Earth to global and environmental change. GCMs differ vastly in their numerical treatment of polar regions because of the so-called \textit{pole-problem} \cite{W2007JMSJ}. The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a, hereafter referred to as \textit{lat-lon} grids). Depending on the numerics, methods exist to suppress this instability, and lat-lon grids may be advantageous for polar processes by representing structures with finer resolution than elsewhere in the computational domain. With the recent trend towards globally uniform unstructured grids, the benefits of lat-lon grids in polar regions may be lost. 
{\color{blue} (WHL: Would it be worth mentioning that quasi-uniform grids also have benefits in polar regions, e.g. by removing elongated grid cells?)}{\color{red} (ARH: I don't think it's worth getting into that here. You'll see later in the manuscript that b/c the cells are narrower in the E-W direction, more cells can fit into the big ablation zone on the western margin.)}{\color{purple}(PHL: another advantage of lat-lon grids is that the coordinate lines are orthogonal!)}
In this study we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids, to understand their impacts on the simulated Arctic climate. We focus on the climate and surface mass balance of the Greenland Ice Sheet.

In the 1970s the pole problem was largely defeated through the adoption of efficient spectral transform methods in GCMs. 
{\color{blue}(WHL: Add refs)}
These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g., horizontal derivatives) in the equation set can be solved exactly. While spectral transform methods are still used today, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems. The pole problem has thus re-emerged in contemporary climate models that use lat-lon grids, and some combination of reduced grids 
{\color{blue}(define 'reduced grids')}
and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}. Polar filters subdue the growth of unstable modes by applying additional damping to the solution over polar regions. This damping reduces the effective resolution in polar regions such that the resolved scales are \textit{approximately} the same everywhere on the grid. We emphasize \textit{approximately} since it's at least conceivable that marginal increases in effective resolution occur over polar region in lat-lon grids, despite polar filtering.

{\color{blue} (WHL: This raises the question: How do lat-lon grids better resolve polar regions at a given nominal global resolution, if the filtering makes them effectively coarser?  Does the damping degrade some processes more than others?)}{\color{red} (ARH - The phrase "approximately the same" doing a lot of heavy lifting here b/c I frankly don't know. The results from this paper suggest that the effective resolution is finer over polar regions, but this is a bit hand-wavy, despite my attempts to quantify it more exactly, and based on the precipitation rates having smaller biases. Also, to be clear, effective resolution is specific to atmospheric features; the land model is free to take advantage of these additional grids cells representing the ablation zone, and they certainly do. Perhaps I can add something to the effect of ``approximately the same everywhere on the grid, although marginal increases in effective resolution over the polar regions cannot be ruled out...")\\

{\color{purple}PHL: These are good points! I think it is useful to separate dynamics from physics in this discussion. Physics is called at more columns on the lat-lon grid as one approaches the poles. There is no "polar filtering" in physics so effectively the physics resolution is much higher than on a the cubed-sphere grid as one approaches the mid-latitudes and poles. Dynamics, on the other hand, filters out smaller scales so effectively it should only retain the 1 degree scale. That said, you will get a more accurate solution since you have more grid-points to represent waves (even though smaller scales are filtered) compared to the cubed-sphere grid. I think that is pretty clear in flow over orography. One can compare the lat-lon grid setup to a spectral model where one keeps viscosity coefficients fixed as one increased horizontal resolution - in such a setup one will get convergence of the dynamics!}}

An alternative approach is to use unstructured grids, which permit quasi-uniform grid spacing globally and eliminates the pole-problem (e.g., Figure~\ref{fig:uni-grids}c).
{\color{blue} (WHL: Add 1 or 2 refs)}
This grid flexibility also allows variable-resolution or regional grid refinement (e.g., Figure~\ref{fig:vr-grids}, hereafter abbreviated as \textit{VR}). In principle, grid refinement over polar regions can make up for any loss of resolution in transitioning away from lat-lon grids (e.g., Figure~\ref{fig:vr-grids}).  However, this refinement comes at the cost of a smaller CFL-limited time step in the refined region. (The CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space.) Unstructured grids also scale more efficiently on parallel systems than lat-lon grids, another advantage as computing power increases over time.

The meteorology and climate of the Arctic are characterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}. For example, while synoptic scale storms are generally well represented at typical GCM resolutions of 1 to 2 degrees \cite{JW2006QJR,IPCCAR5}, mesoscale polar lows are not well resolved at these resolutions. These mesoscale systems are prevalent during winter and produce gale-force winds that can induce large heat and moisture fluxes through the underlying sea-ice/ocean interface. The Arctic also contains the Greenland Ice Sheet (GrIS), which blankets the worlds's largest island.  Many of the processes that control the GrIS annual surface mass balance (SMB; the integrated sum of precipitation and runoff) are only partially resolved at typical GCM resolutions. For example, GrIS precipitation is concentrated at the ice-sheet margins, where steep topographic slopes drive orographic precipitation. GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation. GCMs struggle to resolve the magnitude and extent of these features \cite{P2010CC,VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite<e.g.,>{LETAL2020JAMES}. 
{\color{blue} (WHL: I suggest adding some language explaining why advance is more prevalent than retreat.  E.g.:  ``… of these features.  For example, CESM2 typically underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component.''  Could cite Lofverstrom et al. (2020).)}{\color{red} (ARH: I may delete the reference to this anecdote of advancing in northern GrIS altogether b/c I've been unable to show whether it's related to resolution. I don't have confidence defining a direction of the SMB bias due to resolution because resolution inflates both source and sink terms; the net effect can go either way.)}

This study compares the representation of high-latitude regions using the spectral-element and finite-volume dycores in CESM2.2, as these two dycores treat high latitudes (i.e., the pole problem) in different ways. 
It also serves to document the Arctic-refined grids developed by the authors for CESM2.2. Section~\ref{sec:methods} documents the grids, dycores, and physical parameterizations used in this study, and also describes the experiments, datasets, and evaluation methods. Section~\ref{sec:results} analyzes the results of the experiments, and Section~\ref{sec:conclusions} provides a general discussion and conclusions.

@@

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=60mm]{figs/grid-ARCTIC.pdf}&
         \includegraphics[width=60mm]{figs/grid-ARCTICGRIS.pdf} \\
\end{tabular}
\end{center}
\caption{Spectral-element grid for the VR Arctic grids in this study. Note that this is not the computational grid; each element has $3\times3$ independent grid points.}
\label{fig:vr-grids}
\end{figure}

@@

The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 \cite<CAM;>{CAM63}, supports several different atmospheric dynamical cores. These include dycores on lat-lon grids, such as finite-volume \cite<FV;>{L2004MWR} and Eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models. The EUL dycore is the oldest and least supported dycore in CAM. FV3 is the newest dycore, but was not fully incorporated when this work begin.  This study compares the performance of the SE and FV dycores, omitting the EUL and FV3 dycores.

[phl: maybe mention that FV was used for IPCC simulations; SE in iHESP]

@@

The FV dycore is a hydrostatic model that integrates the equations of motion using a finite-volume discretization on a spherical lat-lon grid \cite{LR1997QJR}. The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to an Eulerian reference grid in the vertical \cite{L2004MWR}, using a hybrid-pressure vertical coordinate. Hyperviscous damping is applied to the divergent modes, while Laplacian damping is applied to momentum in the top few layers, referred to as a \textit{sponge layer} \cite{L2011IJHPC} [phl: as far as I know the Laplacian-like damping is not invoked in lower resolution setups - was implemented for 1/4 applications where the polar night jet became excessively strong]. A polar filter damps computational instability due to the convergence of meridians, allowing a longer time step. It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically in the poleward direction \cite{ST1995GEOS}.

@@

The SE dycore is a hydrostatic model that integrates the equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}. The computational domain is a cubed-sphere grid tiled with quadrilateral elements (see Figure~\ref{fig:vr-grids}). Each element contains a fourth-order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre (GLL) quadrature points. This results in 16 GLL nodal points per element, with 12 of the points lying on the (shared) element boundary. Communication between elements uses the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries to reconcile overlapping nodal values and produce a continuous global basis set.

As with the FV dycore, the dynamics evolve in floating Lagrangian layers that are subsequently mapped to an Eulerian reference grid. A dry mass vertical coordinate was recently implemented for thermodynamic consistency with condensates \cite{LetAl2018JAMES}. The 2D dynamics have no implicit dissipation, and so hyperviscosity operators are applied to all prognostic variables to remove spurious numerical errors \cite{DetAl2012IJHPCA}. Laplacian damping is applied in the sponge layer.

The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements to uniform-resolution grids: (1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing. A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements. (2) The topography boundary conditions need to be smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{gmdd-8-4623-2015} has been modified to scale the smoothing radius by the local element size.

For spectral-element grids with quasi-uniform grid spacing, we use a variant in which tracer advection is computed using the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) \cite<>{LTOUNGK2017MWR}. 
{\color{blue} (WHL: I'm not sure what CSLAM is a variant of, since I don't see a tracer transport scheme described above.  How is tracer transport handled on VR grids?  If it's not too much detail, why does CSLAM not work on VR grids?)} {\color{red} (ARH: fair point. the spectral-element section is lacking in general, so keeping this here as a reminder to explain the spectral-element method is used for tracer transport in ``vanilla" cam-se, and we only support vanilla var-res. PHL is planning on adding a description of enhancements to cam-se introduced in CESM2.2, since this turns out to be the only paper documenting CESM2.2.)}
CSLAM has improved tracer property preservation and accelerated multi-tracer transport. It uses a separate grid from the spectral-element dynamics, dividing each element into $3\times3$ control volumes with quasi-equal area. The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the default SE dycore in which the physics are evaluated at the GLL nodal points \cite{HL2018MWR}.

@@

We evaluate six grids in this study (Table~\ref{tbl:table1}). The FV dycore is run with $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $f09$ and $f19$, respectively (Figure~\ref{fig:uni-grids}a,b). We also run the $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid, referred to as $ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics. We run an additional $1^{\circ}$ CAM-SE-CSLAM grid with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>{HETAL2019JAMES}.

Two VR meshes were developed for the CESM2.2 release to support grid refinement over the Arctic (Figure~\ref{fig:vr-grids}). This paper serves as the official documentation of these grids. The Arctic meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}). The $Arctic$ grid is a $1^{\circ}$ grid with $\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region. The $Arctic-GrIS$ grid is identical to the $Arctic$ grid, but with an additional patch covering the island of Greenland with $\frac{1}{8}^{\circ}$ resolution.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lccccc}
   \hline
   grid name & dycore & $\Delta x_{\mathrm{eq}}$ (km) & $\Delta x_{\mathrm{refine}}$ (km) & $\Delta t_{\mathrm{phys}}$ (s) \\ 
   \hline
   $f19$ & FV & 278 & - &1800 \\
   $f09$ & FV & 139 & - &1800 \\
   $ne30pg2$ & SE-CSLAM & 167 & - & 1800 \\
   $ne30pg3$ & SE-CSLAM & 111 & - & 1800 \\
   $ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 \\
   $Arctic$ & SE & 111 & 28 & 450 \\
   $Arctic-GrIS$ & SE & 111 & 14 & 225 \\
 \hline
 \end{tabular}
  \caption{Grids and dycores used in this study. $\Delta x_{\mathrm{eq}}$ is the average equatorial grid spacing, $\Delta x_{\mathrm{refine}}$ is the grid spacing in the refined region (if applicable), and $\Delta t_{\mathrm{phys}}$ is the physics time step. FV refers to the finite-volume dycore, SE the spectral-element dycore, and SE-CSLAM the spectral-element dycore with CSLAM tracer advection.  We use the $ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$.}
 \label{tbl:table1}
 \end{table*}

The physics time step should depend on grid resolution. Increased horizontal resolution permits faster vertical velocities that reduce characteristic time scales, so the physics time step is reduced to avoid large time truncation errors \cite{HR2018JAMES}. The $ARCTIC$ and $Arctic-GrIS$ grids are run with a $4\times$ and $8\times$ reduction in physics time step relative to the default 1800 s time step for the standard $1^{\circ}$ and $2^{\circ}$ grids (Table~\ref{tbl:table1}).

All grids and dycores in this study use 32 levels in the vertical, with a model top of about 1~hPa or 40~km. However, any grid or dycore can in principle be run with a higher top or finer vertical resolution.

@@

All simulations in this study use the CAM6.3 physical parameterization package \cite<hereafter referred to as the \textit{physics};>{CAM63}. The physics in CAM6 differs from its predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a planetary boundary layer (PBL), shallow convection, and cloud macrophysics scheme. CLUBB is coupled with the MG2 microphysics scheme \cite{MG2}, with prognostic precipitation and classical nucleation theory to represent cloud ice for improved cloud-aerosol interactions. Deep convection is parameterized using a convective quasi-equilibrium, bulk plume mass flux scheme \cite{ZM1995AO,NRJ2008JC} and includes convective momentum transport \cite{RSG2010JAS}. PBL form drag is modeled after \citeA{BBW2004QJRMS}.  Orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale.

Initial simulations with the $ne30pg3$ SE grid produced weaker shortwave cloud forcing relative to the tuned finite-volume dycore. All runs with the SE dycore therefore have two CLUBB parameter changes to provide more realistic cloud forcing and top-of-atmosphere radiation balance. These are CLUBB's $gamma$ parameter, reduced from 0.308 to 0.270, and $c14$, reduced from 2.2 to 1.6. Briefly, $gamma$ scales the width of the sub-grid distribution of vertical velocity, and $c14$ controls the strength of the damping term in the equation for the horizontal component of turbulent kinetic energy. For a thorough explanation of how CLUBB parameters impact the simulated climate, see \cite{GETAL2015JAMES}.

@@

We run all grids and dycores using an identical transient 1979-1998 Atmospheric Model Inter-comparison Project configuration (AMIP), with prescribed monthly SST and sea ice following \citeA{CESMSST}. In CESM terminology, AMIP simuations use the $FHIST$ computational set and run out of the box in CESM2.2.

All grids and dycores simulate the GrIS surface mass balance (SMB). The SMB is the sum of the mass accumulation term,  i.e., precipitation, and the mass sink term, ablation. Ablation is the sum of evaporation/sublimation and liquid runoff, with runoff being a combination of liquid precipitation and snow and ice melt. Not all liquid precipitation runs off the ice sheet; rain can penetrate pore spaces in the snowpack/firn layer and freeze, increasing the snow mass. These processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.
